\section{Project: Hurewicz' theorem}
\label{section:Hurewicz}

Let $R$ be a unitary commutative ring and $G$ be a group. Recall that \emph{group ring} is a free $R$-module which is also a ring. As a free $R$-module, $G$ is a basis for $R[G]$. Thus every element $\alpha\in R[G]$ can be written uniquely as 
a sum of the form 
\[
\alpha=\sum_{g\in G}r_gg,\quad r_g\in R.
\]
Note that the elements of $R$ (more precisely, the elements of the form $r1_G$ for $r\in R$) are central in $R[G]$. 

\begin{exercise}
\label{xca:augmentation_ideal}
    Let $R$ be a commutative ring and $G$ be a group. Let 
    \[
    f\colon R[G] \to R,\quad 
    \sum_{g \in G} r_g g\mapsto \sum_{g \in G} r_g.
    \]
    Prove the following statements:
    \begin{enumerate}
        \item $f$ is a surjective ring
            homomorphism. 
        \item $\{g - 1_G: 1_G\ne g \in G\}$ is a    basis for the free $R$-module $\ker f$.
    \end{enumerate}
\end{exercise}

The kernel of the homomorphism of Exercise \ref{xca:augmentation_ideal} is 
called the \emph{augmentation ideal} of
the group ring. We used this ideal before (e.g. Propositions \ref{pro:KGsemisimple} and Appendix \ref{section:local}). 

\begin{theorem}[Hurewicz]
    \label{thm:Hurewicz}
    \index{Hurewicz' theorem}
    Let $G$ be a group and $I$ be the augmentation ideal of $\Z[G]$. 
    Then $G/[G,G]\simeq I/I^2$ as (abelian) groups. 
\end{theorem}

\begin{proof}
    Let $\varphi\colon G\to I/I^2$, $g\mapsto g-1_G+I^2$. Since $g-1_G\in I$ for all $g\in G$, $\varphi$ is well-defined. The map $\varphi$ is a group homomorphism. Since 
    $(g-1_G)(h-1_G)\in I^2$, 
    \begin{align*}
    \varphi(gh) &= gh-1_G+I^2\\
    &=gh-1_G-(g-1_G)(h-1_G)+I^2+I^2\\
    &=g-1_G+h-1_G+I^2\\
    &=\varphi(g)+\varphi(h)
    \end{align*}
    holds for all $g,h\in G$. 

    Since $[G,G]\subseteq\ker\varphi$, there exists a group homomorphism
    \[
    \overline{\varphi}\colon G/[G,G]\to I/I^2,\quad 
    g[G,G]\mapsto g-1_G+I^2.
    \]
    We claim that $\overline{\varphi}$ is an isomorphism. 
    Let us construct the inverse of $\overline{\varphi}$. Let 
    \[
    \psi\colon I\to G/[G,G],\quad 
    \sum_{g\in G}m_g(g-1_G)\mapsto \left(\prod_{g\in G}g^{m_g}\right)[G,G].
    \]
    Since $G/[G,G]$ is abelian, the map $\psi$ is well-defined, that is
    the order of the factors in $\prod_{g\in G}g^{m_g}$ does not matter. Note that 
    $I^2\subseteq\ker\psi$, as 
    $\{(g-1_G)(h-1_G):g,h\in G\}$ generates the additive group $I^2$ 
    and 
    \begin{align*}
        \psi((g-1_G)(h-1_G))&=\psi( (gh-1_G)-(g-1_G)-(h-1_G))\\
        &=(ghg^{-1}h^{-1})[G,G]\\
        &=[G,G].
    \end{align*}
    Therefore there exists a group homomorphism
    \[
    \overline{\psi}\colon I/I^2\to G/[G,G],\quad 
    \sum_{g\in G}m_g(g-1_G)+I^2\mapsto \left(\prod_{g\in G}g^{m_g}\right)[G,G].
    \]
    A direct calculation shows that $\overline{\psi}$ is the inverse 
    of $\overline{\varphi}$. 
\end{proof}

