\section*{Project: When a group algebra is reduced?}
\label{section:reduced}

\begin{exercise}
    Is the ring $\C[\Z/2]$ reduced? 
\end{exercise}

% Let $G=\langle g:g^2=1\rangle$. If $(a+bg)^2=0$, then
% $(a^2+b^2)+(2ab)g=0$. Thus $a=b=0$. 

\begin{problem}
\label{prob:reduced}
    Let $G$ be a torsion-free group. Is
    $K[G]$ is reduced?
\end{problem}

Problem \ref{prob:reduced} is related to other important
open problems about group algebras 
(e.g. zero-divisors, units, 
indempotents and semisimplicity of group
rings).

\begin{exercise}
\label{xca:reduced_central}
    Prove that idempotents of reduced rings are central. 
\end{exercise}

The previous exercise is used to solve the following problem.

\begin{exercise}
\label{xca:x^3=x}
    Let $R$ be a ring such that $x^3=x$ for all $x\in R$. Prove that
    $R$ is commutative. 
\end{exercise}

Exercise \ref{xca:x^3=x} is hard. 
Even harder is the following exercise:

\begin{exercise}
\label{xca:x^4=x}
    Let $R$ be a ring such that $x^4=x$ for all $x\in R$. Prove
    that $R$ is commutative. 
\end{exercise}

%Other exercises about reduced rings. 

%\begin{exercise}
%\label{xca:reduced}
%    Prove that a ring is reduced if 
%    and only it has no non-zero nilpotent elements. 
%\end{exercise}

\begin{exercise}
\label{xca:reduced=>semiprime}
    Reduced rings are semiprime.
\end{exercise}
 
\begin{theorem}
\label{thm:reduced}
    Let $K$ be a field and $G$ be a group. If $K[G]$
    is reduced, then every finite subgroup of $G$ is normal. 
\end{theorem}

\begin{proof}
    Let $H=\{h_1,\dots,h_n\}$ be a finite normal subgroup of $G$. 
    We claim that $n=|H|$ is invertible in $K$. If $\ch K=0$, this 
    is clear. If $\ch K=p>0$ and $n$ is not invertible in $K$, 
    then $p$ divides $n=|H|$. By Cauchy's theorem, 
    there exists an element $h\in H$ of order $n$, that is 
    $|h|=n$. Since $(1-h)^p=1-h^p=0$ and $K[G]$ is reduced,
    $h=1$, a contradiction. 
    
    Let $\alpha=\frac{1}{n}\sum_{i=1}^nh_i\in K[G]$. Then
    \[
    \alpha^2=\frac{1}{n^2}\sum_{i=1}^n\sum_{j=1}^nh_ih_j
    =\frac{1}{n^2}\sum_{i=1}^nn\alpha=\alpha.
    \]
    Thus $\alpha$ is idempotent. As idempotent 
    element of reduced rings are central (Exercise \ref{xca:reduced_central}), 
    $g\alpha g^{-1}=\alpha$ for all $g\in G$. If $g\in G$, 
    then 
    \[
    \sum_{i=1}^n gh_ig^{-1}=\sum_{i=1}^n h_i.
    \]
    It follows that $H$ is normal in $G$, 
    as for each $i\in\{1,\dots,n\}$ 
    there exists $j\in\{1,\dots,n\}$ such that 
    $gh_ig^{-1}=h_j\in H$. 
\end{proof}

\begin{example}
    If $K$ is a field, then $K[\Sym_3]$ is not reduced. 
    In fact, 
    if 
    \[
    \alpha=(12)+(123)-(132)-(13),
    \]
    then 
    $\alpha^2=0$. 
\end{example}

\begin{exercise}
    Prove that the converse of Theorem \ref{thm:reduced} 
    does not hold. 
\end{exercise}

