\section{Project: Bi-ordered groups}

%Based on Example~\ref{example:k[Z]} we will study 
%some properties of groups. 

Recall that a \emph{total order} is a partial order in which any two elements are comparable. This
means that a total order is a binary relation $\leq$ on some set $X$ 
such that for all $x,y,z\in X$ one has 
the following properties: 
\begin{enumerate}
    \item $x\leq x$.
    \item $x\leq y$ and $y\leq z$ imply $x\leq z$.
    \item $x\leq y$ and $y\leq x$ imply $x=y$.
    \item $x\leq y$ or $y\leq x$. 
\end{enumerate}
A set equipped with a total order is a \emph{totally ordered set}. 

\begin{definition}
	\index{Group!bi-ordered}
	A group $G$ is \emph{bi-ordered} if there exists a total order 
	$<$ in $G$
	such that $x<y$ implies that $xz<yz$ and $zx<zy$ for all $x,y,z\in G$.
\end{definition}

\begin{example}
	The group $\R_{>0}$ of positive real numbers is bi-ordered. 
\end{example}

The multiplicative group $\R\setminus\{0\}$ is not bi-ordered. Why?

\begin{exercise}
	Let $G$ be a bi-ordered group and $x,x_1,y,y_1\in G$. Prove that
	$x<y$ and $x_1<y_1$ imply $xx_1<yy_1$.
\end{exercise}

Clearly, bi-orderability is preserved under taking subgroups. 

\begin{exercise}
	Let $G$ be a bi-ordered group and $g,h\in G$. Prove that $g^n=h^n$
	for some $n>0$ implies $g=h$.
\end{exercise}

The following result goes back to Neumann.

\begin{exercise}
    Let $G$ be a bi-ordered group and $g,h\in G$. Prove that $g^n\in C_G(h)$ if 
    and only if $g\in C_G(h)$.
\end{exercise}

Bi-ordered groups do not behave nicely under extensions:

\begin{exercise}
\label{xca:BO_sequence}
    Let $1\to K\to G\to Q\to 1$ be an exact sequence of groups. Assume that $K$ and $Q$ 
    are bi-ordered. Prove that $G$ is bi-ordered if and only if 
    $x<y$ implies $gxg^{-1}<gyg^{-1}$ for all $x,y\in K$ and $g\in G$. 
\end{exercise}

\begin{definition}
	Let $G$ be a bi-ordered group. The \emph{positive cone} of $G$  
	is the set $P(G)=\{x\in
	G:1<x\}$.
\end{definition}

\begin{proposition}
	\label{pro:biordenableP1}
	Let $G$ be a bi-ordered group and let $P$ be its positive cone. 
	\begin{enumerate}
		\item $P$ is closed under multiplication, i.e. $PP\subseteq P$. 
		\item $G=P\cup P^{-1}\cup \{1\}$ (disjoint union).
		\item $xPx^{-1}=P$ for all $x\in G$.
	\end{enumerate}
\end{proposition}

\begin{proof}
	If $x,y\in P$ and $z\in G$, then, since $1<x$ and $1<y$, it follows that 
	$1<xy$.  Thus $1=z1z^{-1}<zxz^{-1}$. It remains to prove the second claim.  
	If $g\in G$, then $g=1$ or $g>1$ or $g<1$. Note that $g<1$ if and only if 
	$1<g^{-1}$, so the claim follows. 
\end{proof}

The previous proposition admits a converse statement. 

\begin{proposition}
	\label{pro:biordenableP2}
	Let $G$ be a group and $P$ be a subset of $G$ such that 
	$P$ is closed under multiplication, $G=P\cup P^{-1}\cup \{1\}$ (disjoint union) and
	$xPx^{-1}=P$ for all $x\in G$. Let $x<y$ whenever  
	$yx^{-1}\in P$. Then $G$ is bi-ordered with positive
	cone is $P$.
\end{proposition}

\begin{proof}
	Let $x,y\in G$. Since $yx^{-1}\in G$ and $G=P\cup
	P^{-1}\cup\{1\}$ (disjoint union), 
	either $yx^{-1}\in P$ or $xy^{-1}=(yx^{-1})^{-1}\in
	P$ or $yx^{-1}=1$. Thus either $x<y$ or $y<x$ or $x=y$. If $x<y$ and $z\in
	G$, then $zx<zy$, as $(zy)(zx)^{-1}=z(yx^{-1})z^{-1}\in P$ and  
	$zPz^{-1}=P$. Moreover, $xz<yz$ since $(yz)(xz)^{-1}=yx^{-1}\in P$. To prove
	that $P$ is the positive cone of $G$ note that 
	$x1^{-1}=x\in P$ if and only if $1<x$. 
\end{proof}

An important property:

\begin{proposition}
	\label{pro:BOsintorsion}
	Bi-ordered groups are torsion-free.
\end{proposition}

\begin{proof}
	Let $G$ be a bi-ordered group and $g\in G\setminus\{1\}$. 
	If $g>1$, then
	$1<g<g^2<\cdots$. If $g<1$, then $1>g>g^2>\cdots$. Hence $g^n\ne 1$ 
	for all $n\ne 0$. 
\end{proof}

The converse of the previous proposition does not hold. 

\begin{exercise}
Let $G=\langle x,y:yxy^{-1}=x^{-1}\rangle$. 
\begin{enumerate}
    \item Prove that $x$ and $y$ are torsion-free. 
    \item Prove that $G$ is torsion-free. 
    \item Prove that $G\simeq \langle a,b:a^2=b^2\rangle$.
\end{enumerate}
\end{exercise}

% To prove that $x$ is torsion-free, use the group homomorphism $G\to\Z$
% given by $x\mapsto 1$, $y\mapsto 0$. To prove that $y$ is torsion-free
% use the group homomorphism $x\mapsto\begin{pmatrix}-1&0\\0&1\end{pmatrix}$,
% $y\mapsto\begin{pmatrix}1&1\\0&1\end{pmatrix}$. 

% 	We first show that $G\simeq\langle a,b:a^2=b^2\rangle$. For that purpose, 
% 	it is enough to check that 
% 	the map $G\to \langle a,b:a^2=b^2\rangle$, $x\mapsto a$, $y\mapsto ab^{-1}$, 
% 	is a well-defined group homomorphism with inverse
% 	$\langle a,b:a^2=b^2\rangle\to G$, $a\mapsto x$, $b\mapsto y^{-1}x$. 

\begin{example}
	The torsion-free group $G=\langle x,y:yxy^{-1}=x^{-1}\rangle$ is not bi-ordered. 
	If not, let $P$ 
	be the positive cone. If $x\in P$, 
	then $yxy^{-1}=x^{-1}\in P$, a contradiction. Hence $x^{-1}\in P$
	and $x=y^{-1}x^{-1}y\in P$, a contradiction.
\end{example}

Bi-ordered groups provide 
an answer to Kaplansky's unit problem (see Question \ref{question:units}).

\begin{theorem}
	\label{thm:BO}
	Let $G$ be a bi-ordered group. Then $K[G]$ is a domain such that
	only has trivial units. Moreover, if $G$ is non-trivial, 
	then $J(K[G])=\{0\}$. 
\end{theorem}

\begin{proof}
	Let $\alpha,\beta\in K[G]$ be such that  
	\begin{align*}
		\alpha&=\sum_{i=1}^m a_ig_i, && g_1<g_2<\cdots<g_m,&& a_i\ne 0 && \forall i\in\{1,\dots,m\},\\
		\beta&=\sum_{j=1}^n b_jh_j, && h_1<h_2<\cdots<h_n, && b_j\ne 0 && \forall j\in\{1,\dots,n\}.
	\end{align*}
	Then 
	\[
		g_1h_1\leq g_ih_j\leq g_mh_n
	\]
	for all $i,j$. Moreover, $g_1h_1=g_ih_j$ if and only if $i=j=1$. The
	coefficient of $g_1h_1$ in $\alpha\beta$ is $a_1b_1\ne 0$. In particular, 
	$\alpha\beta\ne0$. If $\alpha\beta=\beta\alpha=1$, then the coefficient of
	$g_mh_n$ in $\alpha\beta$ is $a_mb_n$. Hence $m=n=1$ and therefore 
	$\alpha=a_1g_1$ and $\beta=b_1h_1$ with $a_1b_1=b_1a_1=1$ in $K$ and $g_1h_1=1$
	in $G$.
\end{proof}

\begin{theorem}[Levi]
	\label{thm:Levi}
	\index{Levi's theorem}
	Let $A$ be an abelian group. Then $A$ is bi-ordered if and only
	if $A$ is torsion-free.
\end{theorem}

\begin{proof}
	If $A$ is bi-ordered, then $A$ is torsion-free. Let us prove the non-trivial implication, 
	so assume that
	$A$ is torsion-free abelian. Let $\mathcal{S}$ be the class 
	of subsets $P$ of $A$ such that $0\in P$, are closed under 
	the addition of
	$A$ and satisfy the following property: if $x\in P$ and $-x\in P$,
	then $x=0$.
	Clearly, $\mathcal{S}\ne\emptyset$, as 
	$\{0\}\in\mathcal{S}$.  The inclusion turns $\mathcal{S}$ into a partially ordered set  
	and $\bigcup_{i\in I}P_i$ is an upper bound for the chain 
	$\{P_i:i\in I\}$. By Zorn's lemma, 
	$\mathcal{S}$ admits a maximal element $P\in\mathcal{S}$.

	\begin{claim}
		If $x\in A$ is such that $kx\in P$ for some $k>0$, then  $x\in P$.		
	\end{claim}

	Let $Q=\{x\in A:kx\in P\text{ for some 
	$k>0$}\}$. We will show that $Q\in\mathcal{S}$.  Clearly, $0\in Q$. Moreover, $Q$
	is closed under addition, as $k_1x_1\in P$ and $k_2x_2\in P$ imply 
	$k_1k_2(x_1+x_2)\in P$. Let $x\in A$ be such that $x\in Q$ and $-x\in Q$. Thus 
	$kx\in P$ and $l(-x)\in P$ for some $l>0$. Since $klx\in P$ and 
	$kl(-x)\in P$, it follows that $klx=0$, a contradiction since $A$ is torsion-free. 
	Hence $x\in Q\subseteq P$. 

	\begin{claim}
		If $x\in A$ is such that $x\not\in P$, then $-x\in P$. 	
	\end{claim}

	Assume that $-x\not\in P$ and let $P_1=\{y+nx:y\in P,\,n\geq0\}$. We will
	show that  $P_1\in\mathcal{S}$.  Clearly, $0\in P_1$ and $P_1$ is closed under
	addition. If $P_1\not\in S$, there exists 
	\[
		0\ne y_1+n_1x=-(y_2+n_2x),
	\]
	where $y_1,y_2\in P$ and $n_1,n_2\geq0$. Thus $y_1+y_2=-(n_1+n_2)x$. If 
	$n_1=n_2=0$, then $y_1=-y_2\in P$ and $y_1=y_2=0$, so it follows that
	$y_1+n_1x=0$, a contradiction. If $n_1+n_2>0$, then, since 
	\[
		(n_1+n_2)(-x)=y_1+y_2\in P,
	\]
	it follows from the first claim that $-x\in P$, a contradiction. 
	Let us show that $P_1\in\mathcal{S}$. 
	Since $P\subseteq P_1$, the maximality of $P$ implies that 
	$x\in P=P_1$.

	\medskip
	By Proposition~\ref{pro:biordenableP2}, 
	$P^*=P\setminus\{0\}$ is the positive cone of a bi-order in $A$. 
	In fact, $P^*$ is closed under addition, as $x,y\in
	P^*$ implies that $x+y\in P$ and $x+y=0$ implies $x=y=0$, as $x=-y\in P$. Moreover,
	$G=P^*\cup -P^*\cup\{0\}$ (disjoint union), as 
	the second claim states that $x\not\in P^*$ implies 
	$-x\in P$. 
\end{proof}

Our proof of Passman's theorem (see Theorem \ref{thm:Passman}) 
used the fact that the group algebra $K[G]$ of
a torsion-free abelian group $G$ has no non-zero divisors. 

\begin{exercise}
\label{xca:domain_G_abelian}
	Let $A$ be a non-trivial torsion-free abelian group. Prove that $K[A]$ 
	is a domain that only admits trivial units and $J(K[A])=\{0\}$. 
\end{exercise}

% \begin{proof}
% 	Apply Levi's theorem and Theorem~\ref{thm:BO}.
% \end{proof}

% The first one is a variation on Exercise \ref{xca:BO_sequence}.

\begin{exercise}
    Let $N$ be a central subgroup of $G$. If $N$ and $G/N$ are bi-ordered, 
    then $G$ is bi-ordered. Prove with an example that $N$ needs to be central, normal 
    is not enough. 
\end{exercise}

\begin{exercise}
    Let $G$ be a group that admits 
    a sequence 
    \[
    \{1\}=G_0\subseteq G_1\subseteq\cdots\subseteq G_n=G
    \]
    such that
    each $G_k$ is normal in $G_{k+1}$ and each quotient $G_{k+1}/G_k$ is 
    torsion-free abelian. Prove that $G$ is bi-ordered.  
\end{exercise}

\begin{exercise}
    Prove that torsion-free nilpotent groups are bi-ordered. 
\end{exercise}

